### 2010-05-05 paper/paper.tex
Add empty paper

\documentclass{acmsiggraph}

\usepackage[scaled=.92]{helvet}
\usepackage{times}
\usepackage{parskip}
\onlineid{0}

\title{Plotting the Single-Electron Solution to Schr\"{o}dinger's Equation with OpenCL}

\author{Tim Horton\thanks{e-mail: hortot2@rpi.edu}\\Rensselaer Polytechnic Institute}

\begin{document}

\maketitle

\section{Abstract}

Stuff

\section{Software and Hardware}

Stuff

% \bibliographystyle{acmsiggraph}
% \nocite{*}
% \bibliography{proposal}

\end{document}

### 2010-05-05 paper/paper.tex
Add image

@@
 \usepackage{parskip}
+\usepackage{graphicx}
 \onlineid{0}
@@
 
+\includegraphics[width=84.5mm]{320.png}
+
 \section{Abstract}
 
-Stuff
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Praesent ac risus eu ante hendrerit pharetra.
+Cras vulputate scelerisque gravida.
+Nullam vestibulum pellentesque mauris non auctor.
+Quisque at orci nec lorem molestie lacinia eu vitae ligula.
+Morbi scelerisque ipsum quis leo elementum ultricies.
+Sed sem nulla, pellentesque quis volutpat blandit, porttitor at quam.
+Donec pharetra elementum magna nec volutpat.
+Fusce condimentum aliquam congue.
+Cras faucibus egestas magna, et gravida tellus suscipit sit amet.
+Mauris aliquet auctor ultrices.
+
+Donec est mauris, pretium sit amet dictum eu, porttitor ac arcu.
+Donec euismod placerat lacus, nec malesuada turpis porttitor at.
+Donec bibendum felis ornare tortor ultricies rhoncus.
+In hac habitasse platea dictumst.
+In in nisi eros, eget eleifend eros.
+In sodales eleifend est vitae rhoncus.
+Maecenas condimentum cursus ante, viverra semper est hendrerit at.
+Duis tempor convallis arcu, id elementum dolor malesuada nec.
+Curabitur volutpat dapibus orci in pellentesque.
+Curabitur eleifend turpis in massa pharetra commodo.
+Mauris condimentum nisi vel metus faucibus congue a in urna.
+Nulla facilisi.
+In hac habitasse platea dictumst.
+Phasellus eget augue tellus.
+Aenean tristique nisl non arcu mollis a rutrum purus laoreet.
+Integer tempor gravida elit ac congue.
+Mauris nisi purus, auctor et pretium ac, imperdiet eu nulla.
+Nunc non diam ac diam mollis malesuada auctor sed augue.
+Sed in pretium felis.
+
+In sed elit vel leo lacinia rhoncus vel at orci.
+Phasellus eu lacus dapibus nibh imperdiet volutpat quis vel dui.
+Nunc pulvinar laoreet tortor, nec congue dui consectetur eu.
+Pellentesque congue risus quis nulla aliquam a accumsan libero porttitor.
+Mauris eros erat, eleifend in fermentum congue, dapibus nec urna.
+In eget risus porttitor magna malesuada condimentum ac vitae justo.
+Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
+Ut ornare vestibulum nisl et egestas.
+Donec quis nunc nec neque adipiscing adipiscing.
+Quisque eget metus dui, ut vulputate est.
+Quisque luctus condimentum lectus, sit amet semper est aliquam eget.
+In risus lacus, malesuada eu eleifend gravida, porta sit amet elit.
+Vivamus neque ipsum, faucibus id porttitor in, pharetra ut lorem.
+Donec justo lacus, dictum id facilisis eu, sodales in lacus.
+Cras porta, ante a vestibulum commodo, turpis mauris pellentesque tellus, vitae iaculis velit magna id risus.
+Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
+Suspendisse condimentum facilisis enim, a mattis ipsum aliquam in.
+
+Nam venenatis mauris a mauris sagittis varius.
+In hac habitasse platea dictumst.
+Vestibulum a nisl eros.
+Vestibulum rutrum scelerisque massa nec commodo.
+In quam tellus, feugiat et tempor nec, vestibulum eget purus.
+Nulla purus libero, commodo id fringilla in, tristique ut enim.
+Nulla congue dictum ultrices.
+Proin a ipsum nisl, eu pretium nunc.
+Ut egestas suscipit lacus, a consectetur ante suscipit vel.
+Nam interdum lacinia est dapibus tempus.
+Ut tincidunt magna eget eros vulputate id dapibus odio sollicitudin.
+Cras quam lectus, elementum id auctor ac, volutpat ac libero.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Nulla suscipit rutrum interdum.
+
+Cras ultricies, justo eu tempus tempus, ligula augue pulvinar urna, ut convallis enim justo vel purus.
+Donec ornare, felis eget blandit luctus, sem diam rutrum libero, vel vehicula sapien dui non enim.
+Quisque porttitor interdum enim.
+Integer sit amet sem sapien, ut pharetra nibh.
+Duis at placerat tellus.
+Donec venenatis erat vestibulum lectus consectetur vel sodales purus rhoncus.
+Sed dapibus odio commodo lectus egestas suscipit.
+Proin placerat ullamcorper mauris nec luctus.
+Etiam rhoncus, nunc vitae eleifend hendrerit, orci massa mollis neque, eget congue lectus mi non orci.
+Vivamus in neque a elit convallis fringilla.
 

### 2010-05-05 hydrogen.py
Fix stupid mistake

@@ -63,7 +63,7 @@ def exportImage(output):
 def benchmark(skip, params=(3,2,0)):
     """Time computation of the image at various different resolutions"""
     for res in range(100,1010,skip):
-        minDuration = 0
+        minDuration = 100000000
 
         for itr in range(5):
             (output, duration) = renderOrbitals(params, res)

### 2010-05-05 paper/build.sh
Add bibliography

@@ -1,0 +1,1 @@
+pdflatex paper.tex && bibtex paper && pdflatex paper.tex && open ./paper.pdf

### 2010-05-05 paper/paper.tex
Add bibliography
Add some stuff to bibliography
Hardware specs, some manner of abstract

--- 2010-05-05 paper/paper.tex
@@
 \usepackage{parskip}
+\usepackage{graphicx}
+\usepackage{footmisc}
 \onlineid{0}
@@
 
+\includegraphics[width=84.5mm]{320.png}
+
 \section{Abstract}
 
-Stuff
+The realm of physics provides many readily parallelizable algorithms --- often involving the simple evaluation of a function at an enormous number of points in space --- which also happen to produce attractive visualizations.
+One such visualization is that of the probability distribution of the single electron within a hydrogen atom (or He$^+$, Li$^{2+}$, and so on).
+By evaluating Schr\"{o}dinger's equation at millions of points, one can construct a representation of the likely location of the electron --- the orbital cloud, if you will.
+Luckily, the evaluation of this function at a particular point is entirely independent of nearby points, so it is effectively perfect for parallelization.
+We will use OpenCL to develop and benchmark (on a varied array of hardware, from dated CPUs to very modern GPUs) an implementation of this algorithm which will produce a simple --- but attractive --- image of the atomic orbital of the single-electron hydrogen atom.
+
+\section{Introduction}
+
+asdfasdfadsf asdf adsf asdf asdf adfasdf adsf adf asfd
+
+asdfasdfadsf asdf adsf asdf asdf adfasdf adsf adf asfd
+
+asdfasdfadsf asdf adsf asdf asdf adfasdf adsf adf asfd
+
+asdfasdfadsf asdf adsf asdf asdf adfasdf adsf adf asfd
+
+\section{Physics}
+
+\section{Implementation}
 
@@
 
-Stuff
+Benchmarks will be performed on a number of different OpenCL compute devices across a few different computers:
 
-% \bibliographystyle{acmsiggraph}
-% \nocite{*}
-% \bibliography{proposal}
+\begin{itemize}
+
+\item ATI Radeon 4890, 800$\times$850MHz, 1GB VRAM, 250\$\footnote{All hardware prices listed are approximate launch prices.\label{fn:prices}}
+
+\item Intel Core i7 620M, 2$\times$3333MHz, 4GB RAM, 332\$\footref{fn:prices}
+
+\item Intel Core 2 Quad Q6600, 4$\times$3000MHz, 4GB RAM, 851\$\footref{fn:prices}
+
+\item Intel Atom N270, 1$\times$1600MHz, 2GB RAM, 44\$\footref{fn:prices}
+
+\end{itemize}
+
+\section{Results}
+
+\subsection{Large- vs. Small-scale Parallelism}
+
+\subsection{Strong Scaling}
+
+\subsection{Cost Effectiveness of Hardware}
+
+\section{Future Work}
+
+\section{Conclusion}
+
+\bibliographystyle{acmsiggraph}
+\nocite{*}
+\bibliography{paper}
 

### 2010-05-05 paper/paper.tex
OS notes, GPU non-general-purpose

@@
 
-\section{Software and Hardware}
+\section{Hardware}
 
@@
 \end{itemize}
+
+These machines run a variety of different operating systems (including Mac OS X, Windows, and Linux).
+Comparisons made later in this paper assume that each OS and the drivers and OpenCL implementation used within them are created equally; this is only somewhat reasonable, and should be kept in mind when interpreting results.
+
+Also, it should be noted that while each core on a given CPU could potentially be evaluating samples in parallel, a GPU's cores are much more restricted (less general-purpose) and must work in small groups to accomplish their work.
+Therefore, given the kernel used for this project, the 4890 listed above only has 50 compute units (16 stream processors work together to compute a single sample).
 

### 2010-05-05 paper/paper.tex
Code note

@@
 \usepackage{footmisc}
+\usepackage{url}
 \onlineid{0}
@@
 
-\item ATI Radeon 4890, 800$\times$850MHz, 1GB VRAM, 250\$\footnote{All hardware prices listed are approximate launch prices.\label{fn:prices}}
+\item ATI Radeon 4890, 800$\times$850MHz, 1GB VRAM, 250\$\footnote{All hardware prices listed are approximate launch prices.
+Prices of older hardware, especially the Core 2 Quad, have dropped significantly since introduction.\label{fn:prices}}
 
@@
 
+\section{Code}
+
+All of the code developed for this project is available under the two-clause BSD license, and is hosted on GitHub:
+
+\url{http://github.com/hortont424/orbitals}
+
+\url{git://github.com/hortont424/orbitals.git}
+
 \bibliographystyle{acmsiggraph}

### 2010-05-05 paper/paper.tex
Some math
Split some math
Strong-scalability statistics (on 4890), lots of equations

@@
 
+\begin{figure}
     \includegraphics[width=84.5mm]{320.png}
+    \caption{The 3-2-0 orbital of the hydrogen atom}
+\end{figure}
 
@@
 
+\subsection{Schr\"{o}dinger's Equation}
+
+From \cite{quantumBook}, we find the single-electron solution to Schr\"{o}dinger's equation:
+
+\begin{equation}\label{psi}
+\psi\left(n, l, m\right)=\psi_c\left(n, l\right)
+\mathit{e}^{-r/na}
+\left(\frac{2r}{na}\right)^l
+\left[L_{n-l-1}^{2l+1}
+    \left(\frac{2r}{na}\right)\right]
+Y_l^m\left(\theta,\phi\right)
+\end{equation}
+
+\begin{equation}\label{psiConstant}
+\psi_c\left(n, l\right)=\sqrt{\left(\frac{2}{na}\right)^3
+    \frac{\left(n-l-1\right)!}{2n\left[\left(n+l\right)!\right]^3}}
+\end{equation}
+
+One will notice that $\psi_c$ is independent of the coordinates of evaluation, $\left(\theta, \phi, r\right)$, so it --- as an obvious optimization --- can be computed once and cached for the entire image.
+
+Also, $\psi$ depends on two external functions: $Y$, the spherical harmonic equation (equation \ref{yFunc}), and $L$, the Laguerre polynomial (equation \ref{laguerre}).
+
+\subsection{Spherical Harmonics\label{sphericalHarmonics}}
+
+\begin{equation}\label{yFunc}
+Y_l^m\left(\theta,\phi\right)=\epsilon
+\sqrt{\frac{\left(2l+1\right)}{4\pi}
+    \frac{\left(l-\left|m\right|\right)!}{\left(l+\left|m\right|\right)!}}
+\mathit{e}^{{\rm i}m\phi}
+P^m_l\left(\cos\theta\right)
+\end{equation}
+
+$Y$ depends on an additional external function, $P$, the Legendre polynomial (equation \ref{legendre}).
+
+\subsection{Laguerre and Legendre\label{laguerreLegendre}}
+
+Key to the evaluation of Schr\"{o}dinger's equation are the Laguerre and Legendre polynomials. Unfortunately,
+
+\begin{equation}\label{laguerre}
+L^\alpha_n\left(x\right)=\frac{x^{-\alpha}e^x}{n!}\frac{d^n}{dx^n}\left(e^{-x}x^{n+\alpha}\right)
+\end{equation}
+
+\begin{equation}\label{legendre}
+P=
+\end{equation}
+
 \section{Implementation}
@@
 
-Benchmarks will be performed on a number of different OpenCL compute devices across a few different computers:
+Benchmarks will be performed on a number of different computation devices across a few different computers:
 
@@
 
+In order to measure the strong scalability of our implementation, we fix the problem size and vary the number of processing units used to compute it.
+For this benchmark, we will fix the image at $400\times400$ pixels, requiring $320,000,000$ total samples, and we will vary the number of cores of the 4890 that we use.
+
 \subsection{Cost Effectiveness of Hardware}

### 2010-05-05 paper/paper.tex
Lots more writing about formulas and cleaup and stuff

@@
 \usepackage{footmisc}
+\usepackage{amsmath}
 \usepackage{url}
@@
 \begin{equation}\label{psi}
-\psi\left(n, l, m\right)=\psi_c\left(n, l\right)
-\mathit{e}^{-r/na}
+\psi\left(n, l, m\right)=\psi_c
+\left(\mathit{e}^{-r/na}\right)
 \left(\frac{2r}{na}\right)^l
@@
     \left(\frac{2r}{na}\right)\right]
-Y_l^m\left(\theta,\phi\right)
+Y_l^m\left(\theta,\phi\right),
 \end{equation}
@@
 \begin{equation}\label{psiConstant}
-\psi_c\left(n, l\right)=\sqrt{\left(\frac{2}{na}\right)^3
-    \frac{\left(n-l-1\right)!}{2n\left[\left(n+l\right)!\right]^3}}
+\psi_c=\sqrt{\left(\frac{2}{na}\right)^3
+    \frac{\left(n-l-1\right)!}{2n\left[\left(n+l\right)!\right]^3}}.
 \end{equation}
@@
 
+Gotta write a blurb here:
+
 \begin{equation}\label{yFunc}
@@
 \mathit{e}^{{\rm i}m\phi}
-P^m_l\left(\cos\theta\right)
+P^m_l\left(\cos\theta\right),
+\end{equation}
+
+\begin{equation}\label{yEpsilon}
+\epsilon=\begin{cases}
+\left(-1\right)^m & \text{$m\ge0$} \\
+1 & \text{$m<0$}
+\end{cases}.
 \end{equation}
@@
 
-Key to the evaluation of Schr\"{o}dinger's equation are the Laguerre and Legendre polynomials. Unfortunately,
+Key to the evaluation of Schr\"{o}dinger's equation are the Laguerre and Legendre polynomials.
+Unfortunately, the generation of both requires symbolically solving differential equations.
+Since OpenCL (rightfully) lacks a symbolic differential equation solver (and the implementation of such a program is far outside of the scope of this project), we have added to our kernel a simple table of the first few polynomials, computed with Mathematica.
+This limits the range of input $n, l, m$ parameters which can be used with this program, but extending the tables is very simple and requires only patience.
+
+The generating differential equations for these polynomials are as follows:
 
 \begin{equation}\label{laguerre}
-L^\alpha_n\left(x\right)=\frac{x^{-\alpha}e^x}{n!}\frac{d^n}{dx^n}\left(e^{-x}x^{n+\alpha}\right)
+L^\alpha_n\left(x\right)=\frac{x^{-\alpha}e^x}{n!}\frac{d^n}{dx^n}\left(e^{-x}x^{n+\alpha}\right),
 \end{equation}
@@
 \begin{equation}\label{legendre}
-P=
+P^u_v\left(z\right)=\frac{\left(1+z\right)^{\mu/2}}{\left(1-z\right)^{\mu/2}}
+\tilde{F}_{2,1}
+\left(-v,v+1;1-\mu;\frac{1-z}{2}\right),
 \end{equation}
+
+where $\tilde{F}$ is Gauss' hypergeometric function, which we don't care about since we're simply allowing Mathematica to evaluate it and using the comparatively simple symbolic results in our table.
 
@@
 
+MAKE SURE TO CITE MATHWORLD
+
 \end{document}

### 2010-05-05 paper/paper.tex
More formulas and citations
Plots!

@@
 
+\begin{figure}
+    \includegraphics[width=84.5mm]{strongPlotOne.pdf}
+    \caption{Core Count vs. Runtime on the 4890 with a fixed resolution ($400\times400$)}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{strongPlotTwo.pdf}
+    \caption{Core Count vs.
+Speedup (measured against a single-core run) on the 4890 with a fixed resolution ($400\times400$); the orange line is the theoretical maximum speedup, given linear scaling and 50 execution units (800 cores / 16 cores per unit); the red line is a linear (1:1) speedup curve}
+\end{figure}
+
 \subsection{Cost Effectiveness of Hardware}
@@
 
-MAKE SURE TO CITE MATHWORLD
-
 \end{document}

### 2010-05-05 paper/paper.tex
Lots more stats

--- 2010-05-05 paper/paper.tex
@@
 Unfortunately, the generation of both requires symbolically solving differential equations.
-Since OpenCL (rightfully) lacks a symbolic differential equation solver (and the implementation of such a program is far outside of the scope of this project), we have added to our kernel a simple table of the first few polynomials, computed with Mathematica.
+\cite{legendreCite} Since OpenCL (rightfully) lacks a symbolic differential equation solver (and the implementation of such a program is far outside of the scope of this project), we have added to our kernel a simple table of the first few polynomials, computed with Mathematica.
 This limits the range of input $n, l, m$ parameters which can be used with this program, but extending the tables is very simple and requires only patience.
@@
 
+Our implementation is twofold: an OpenCL kernel (written in the C-like OpenCL kernel language) which performs the evaluation of all of the Schr\"{o}dinger's equation samples required for a single pixel of the output image, and a Python script which parses command line options, uses PyOpenCL to load and run the kernel and PIL to create the output image, and performs timed benchmarks.
+
+The decision was made during the implementation phase to redefine the Bohr radius so that $a=1$; this was done in order to keep intermediate numbers within the range of single-precision floating point values, as OpenCL doesn't currently support native double-precision math on any of the GPUs available for benchmarking.
+Since we're only using this software to generate visualizations, this doesn't affect the output; it simply scales the distances away from the atomic level.
+
 \section{Hardware}
@@
 
+\subsection{GPUs}
+
 \begin{itemize}
 
-\item ATI Radeon 4890, 800$\times$850MHz, 1GB VRAM, 250\$\footnote{All hardware prices listed are approximate launch prices.
+\item ATI Radeon 4890, 800$\times$850MHz, 1GB, 250\$\footnote{All hardware prices listed are approximate launch prices.
 Prices of older hardware, especially the Core 2 Quad, have dropped significantly since introduction.\label{fn:prices}}
 
-\item Intel Core i7 620M, 2$\times$3333MHz, 4GB RAM, 332\$\footref{fn:prices}
+\item NVIDIA GeForce 8600M GT, 32$\times$475MHz, 512MB, ???\$\footref{fn:prices}
 
-\item Intel Core 2 Quad Q6600, 4$\times$3000MHz, 4GB RAM, 851\$\footref{fn:prices}
+\end{itemize}
 
-\item Intel Atom N270, 1$\times$1600MHz, 2GB RAM, 44\$\footref{fn:prices}
+\subsection{CPUs}
+
+\begin{itemize}
+
+\item Intel Core i7 620M, 2$\times$3333MHz, 4GB, 332\$\footref{fn:prices}
+
+\item Intel Core 2 Quad Q6600, 4$\times$3000MHz, 4GB, 851\$\footref{fn:prices}
+
+\item Intel Core 2 Duo E7200, 2$\times$2530MHz, 8GB, 133\$\footref{fn:prices}
 
@@
 
+\begin{figure}
+    \includegraphics[width=84.5mm]{runtimePlot.pdf}
+    \caption{Runtime vs. Resolution on All Hardware}
+    \label{fig:runtimePlot}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{speedupPlot.pdf}
+    \caption{Runtime vs. Resolution on All Hardware}
+    \label{fig:runtimePlot}
+\end{figure}
+
 \subsection{Strong Scaling}
@@
 For this benchmark, we will fix the image at $400\times400$ pixels, requiring $320,000,000$ total samples, and we will vary the number of cores of the 4890 that we use.
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{strongPlotOne.pdf}
+    \caption{Core Count vs. Runtime on the 4890 with a fixed resolution ($400\times400$)}
+    \label{fig:strongPlotOne}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{strongPlotTwo.pdf}
+    \caption{Core Count vs.
+Speedup (measured against the single-core run) on the 4890 with a fixed resolution ($400\times400$); the orange line is the theoretical maximum speedup, given linear scaling and 50 execution units (800 cores / 16 cores per unit); the red line is a linear speedup curve}
+    \label{fig:strongPlotTwo}
+\end{figure}
+
+One can see in figure \ref{fig:strongPlotTwo} that this implementation is strongly scalable; the speedup is almost perfectly linear as the core count increases.
+It should be noted that since the 4890 only has a maximum of 50 compute units, the linear increase is only maintained until we reach 50 cores; after that, the performance gains plateau as they should.
 
@@
 
+The code has only been tested with the Apple and ATI OpenCL compilers, but should work with few to no changes on NVIDIA's SDK.
+
 \bibliographystyle{acmsiggraph}

### 2010-05-05 paper/paper.tex
More figures, more words

@@
 
-\subsection{Large- vs. Small-scale Parallelism}
+\subsection{Strong Scaling}
 
-\begin{figure}
-    \includegraphics[width=84.5mm]{runtimePlot.pdf}
-    \caption{Runtime vs. Resolution on All Hardware}
-    \label{fig:runtimePlot}
-\end{figure}
-
-\begin{figure}
-    \includegraphics[width=84.5mm]{speedupPlot.pdf}
-    \caption{Runtime vs. Resolution on All Hardware}
-    \label{fig:runtimePlot}
-\end{figure}
-
-\subsection{Strong Scaling}
+\label{strongScaling}
 
@@
     \includegraphics[width=84.5mm]{strongPlotOne.pdf}
-    \caption{Core Count vs. Runtime on the 4890 with a fixed resolution ($400\times400$)}
+    \caption{Core Count vs. Runtime on the 4890 with a fixed resolution ($400\times400$); smaller values are better}
     \label{fig:strongPlotOne}
@@
     \caption{Core Count vs.
-Speedup (measured against the single-core run) on the 4890 with a fixed resolution ($400\times400$); the orange line is the theoretical maximum speedup, given linear scaling and 50 execution units (800 cores / 16 cores per unit); the red line is a linear speedup curve}
+Speedup (measured against the single-core run) on the 4890 with a fixed resolution ($400\times400$); the orange line is the theoretical maximum speedup, given linear scaling and 50 execution units (800 cores / 16 cores per unit); the red line is a linear speedup curve; larger values are better}
     \label{fig:strongPlotTwo}
@@
 It should be noted that since the 4890 only has a maximum of 50 compute units, the linear increase is only maintained until we reach 50 cores; after that, the performance gains plateau as they should.
+
+\subsection{Large- vs. Small-scale Parallelism}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{runtimePlot.pdf}
+    \caption{Runtime vs. Resolution on all hardware; smaller values are better}
+    \label{fig:runtimePlot}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{speedupPlot.pdf}
+    \caption{Speedup of fastest GPU over fastest CPU as resolution increases; larger values are better}
+    \label{fig:speedupPlot}
+\end{figure}
+
+Figure \ref{fig:speedupPlot} shows that our implementation converges to an approximately 14.7x speedup (between the fastest of each class of hardware, namely, the 4890 and Core 2 Quad) as the resolution of the output image increases.
+Noise in the speedup values with smaller images is likely due to measurement error and the overhead and unpredictability involved when copying data to/from video memory.
+
+It should be noted that these speedup values are across different pieces of hardware at different core clock speeds, and, as such, do not represent the actual parallel speedup gained (for that, look at section \ref{strongScaling}).
+This is, instead, a representation of the performance potentially gained by moving computation to the GPU.
 

### 2010-05-05 paper/paper.tex
OS notes

@@
 \item ATI Radeon 4890, 800$\times$850MHz, 1GB, 250\$\footnote{All hardware prices listed are approximate launch prices.
-Prices of older hardware, especially the Core 2 Quad, have dropped significantly since introduction.\label{fn:prices}}
+Prices of older hardware, especially the Core 2 Quad, have dropped significantly since introduction.\label{fn:prices}}\footnote{Tested on Windows, with ATI Stream SDK\label{fn:windows}}
 
-\item NVIDIA GeForce 8600M GT, 32$\times$475MHz, 512MB, ???\$\footref{fn:prices}
+\item NVIDIA GeForce 8600M GT, 32$\times$475MHz, 512MB, ???\$\footref{fn:prices}\footnote{Tested on Mac OS X, with Apple OpenCL\label{fn:osx}}
 
@@
 
-\item Intel Core i7 620M, 2$\times$3333MHz, 4GB, 332\$\footref{fn:prices}
+\item Intel Core i7 620M, 2$\times$3333MHz, 4GB, 332\$\footref{fn:prices}\footref{fn:osx}
 
-\item Intel Core 2 Quad Q6600, 4$\times$3000MHz, 4GB, 851\$\footref{fn:prices}
+\item Intel Core 2 Quad Q6600, 4$\times$3000MHz, 4GB, 851\$\footref{fn:prices}\footnote{Tested on Linux, with ATI Stream SDK\label{fn:linux}}
 
-\item Intel Core 2 Duo E7200, 2$\times$2530MHz, 8GB, 133\$\footref{fn:prices}
+\item Intel Core 2 Duo E7200, 2$\times$2530MHz, 8GB, 133\$\footref{fn:prices}\footref{fn:osx}
 

### 2010-05-05 paper/paper.tex
more clarity about a thing
Remove various things

@@
 \psi\left(n, l, m\right)=\psi_c
-\left(\mathit{e}^{-r/na}\right)
-\left(\frac{2r}{na}\right)^l
+\left(\mathit{e}^{-r/n}\right)
+\left(\frac{2r}{n}\right)^l
 \left[L_{n-l-1}^{2l+1}
-    \left(\frac{2r}{na}\right)\right]
+    \left(\frac{2r}{n}\right)\right]
 Y_l^m\left(\theta,\phi\right),
@@
 \begin{equation}\label{psiConstant}
-\psi_c=\sqrt{\left(\frac{2}{na}\right)^3
+\psi_c=\sqrt{\left(\frac{2}{n}\right)^3
     \frac{\left(n-l-1\right)!}{2n\left[\left(n+l\right)!\right]^3}}.
@@
 
+\subsection{Overview}
+
 Our implementation is twofold: an OpenCL kernel (written in the C-like OpenCL kernel language) which performs the evaluation of all of the Schr\"{o}dinger's equation samples required for a single pixel of the output image, and a Python script which parses command line options, uses PyOpenCL to load and run the kernel and PIL to create the output image, and performs timed benchmarks.
+
+\subsection{Simplifications}
 
@@
 Since we're only using this software to generate visualizations, this doesn't affect the output; it simply scales the distances away from the atomic level.
+The equations above have $a=1$ already substituted in, for brevity.
+
+To keep the implementation of this algorithm within the scope of this project, we also decided to use a fixed-camera orthogonal projection.
+This worked to significantly simplify the process of iteration over all of the samples, as with this restriction, the program simply has to iterate over all of the pixels in the image, evaluating $n$ points in the $z$ dimension for each pixel, without worrying about complex transformations between coordinate systems.
 
@@
 
+As one can see in \ref{fig:runtimePlot}, both GPUs significantly outperform all of the CPUs
+
 Figure \ref{fig:speedupPlot} shows that our implementation converges to an approximately 14.7x speedup (between the fastest of each class of hardware, namely, the 4890 and Core 2 Quad) as the resolution of the output image increases.

### 2010-05-06 paper/paper.tex
More data, more plots, etc.

@@
 \section{Physics}
+
+All of the equations below are borrowed from \cite{quantumBook}, but have been slightly modified to fit our purposes, as discussed in section \ref{simpSection}.
 
@@
 
+\label{simpSection}
+
 The decision was made during the implementation phase to redefine the Bohr radius so that $a=1$; this was done in order to keep intermediate numbers within the range of single-precision floating point values, as OpenCL doesn't currently support native double-precision math on any of the GPUs available for benchmarking.
@@
 To keep the implementation of this algorithm within the scope of this project, we also decided to use a fixed-camera orthogonal projection.
-This worked to significantly simplify the process of iteration over all of the samples, as with this restriction, the program simply has to iterate over all of the pixels in the image, evaluating $n$ points in the $z$ dimension for each pixel, without worrying about complex transformations between coordinate systems.
+This worked to significantly simplify the process of iteration over all of the samples, as with this restriction, the program simply has to iterate over all of the pixels in the image, evaluating many points in the $z$ dimension for each pixel, without worrying about complex transformations between coordinate systems.
+
+\subsection{Evaluation}
+
+For each pixel in the image, we spawn an instance of our OpenCL kernel which evaluates 2000 samples in the $z$ dimension.
+Each sample represents the value of the density function, $\psi$, at that point in space (the cartesian coordinates of the image have to be converted into spherical coordinates in order to evaluate $\psi$).
+All 2000 samples are summed and stored into their respective pixel in the image buffer.
+
+Once computation has completed, the buffer is copied back from video memory to system memory (or, if the program was running on the CPU, this step is skipped --- this is the primary overhead incurred by the use of GPU computation).
+There, pixel values are scaled linearly so that the brightest pixel is 100\% white.
 
@@
 
-\subsection{GPUs}
+\subsection{GPU}
 
@@
 Prices of older hardware, especially the Core 2 Quad, have dropped significantly since introduction.\label{fn:prices}}\footnote{Tested on Windows, with ATI Stream SDK\label{fn:windows}}
-
-\item NVIDIA GeForce 8600M GT, 32$\times$475MHz, 512MB, ???\$\footref{fn:prices}\footnote{Tested on Mac OS X, with Apple OpenCL\label{fn:osx}}
 
@@
 
-\item Intel Core i7 620M, 2$\times$3333MHz, 4GB, 332\$\footref{fn:prices}\footref{fn:osx}
+\item Intel Core i7 620M, 2$\times$3333MHz, 4GB, 332\$\footref{fn:prices}\footnote{Tested on Mac OS X, with Apple OpenCL\label{fn:osx}}
 
@@
     \includegraphics[width=84.5mm]{speedupPlot.pdf}
-    \caption{Speedup of fastest GPU over fastest CPU as resolution increases; larger values are better}
+    \caption{Speedup of fastest GPU (4890) over fastest CPU (Core 2 Quad) as resolution increases; larger values are better}
     \label{fig:speedupPlot}
@@
 
-As one can see in \ref{fig:runtimePlot}, both GPUs significantly outperform all of the CPUs
+As one can see in figure \ref{fig:runtimePlot}, the GPU significantly outperforms all of the CPUs.
+This is not surprising, as it can perform more than an order of magnitude more parallel computations, but it does validate the premise of this experiment.
 
@@
 \subsection{Cost Effectiveness of Hardware}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{dollarPlot.pdf}
+    \caption{Samples per second per dollar with a fixed resolution ($500\times500$); larger values are better}
+    \label{fig:costEffectiveness}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{transistorPlot.pdf}
+    \caption{Samples per second per million transistors with a fixed resolution ($500\times500$); larger values are better}
+    \label{fig:transistorPlot}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{wattPlot.pdf}
+    \caption{Samples per second per heavy-load watt with a fixed resolution ($500\times500$); larger values are better}
+    \label{fig:wattPlot}
+\end{figure}
 

### 2010-05-06 paper/paper.tex
Update some plots

@@
 
+After arguments have been parsed and an OpenCL context has been created, we create a buffer large enough to store the entirety of the resultant image, and also compute the constant term of $\psi$, both of which are passed to each instance of the kernel.
+
 For each pixel in the image, we spawn an instance of our OpenCL kernel which evaluates 2000 samples in the $z$ dimension.
@@
 There, pixel values are scaled linearly so that the brightest pixel is 100\% white.
+
+\subsection{Complex Math}
+
+The OpenCL specification unfortunately currently does not include complex math primitives, which are necessary for the evaluation of $\psi$.
+It does reserve the {\bf complex} keyword, which suggests that perhaps support for something similar to GCC's complex type is coming in a future version of OpenCL, which would put it on ground more similar to CUDA.
+
+In order to solve this problem, we implemented a small library of complex math functions which make use of the OpenCL {\bf float2} type to store complex numbers.
+This library includes various complex number constructors, as well as exponentiation, multiplication, square root, the exponential function, and conjugation.
+These functions are used extensively within the evaluation of $\psi$.
 
@@
     \includegraphics[width=84.5mm]{strongPlotOne.pdf}
-    \caption{Core Count vs. Runtime on the 4890 with a fixed resolution ($400\times400$); smaller values are better}
+    \caption{Runtime as core count increases on the 4890 with a fixed resolution ($400\times400$); smaller values are better}
     \label{fig:strongPlotOne}
@@
     \includegraphics[width=84.5mm]{strongPlotTwo.pdf}
-    \caption{Core Count vs.
-Speedup (measured against the single-core run) on the 4890 with a fixed resolution ($400\times400$); the orange line is the theoretical maximum speedup, given linear scaling and 50 execution units (800 cores / 16 cores per unit); the red line is a linear speedup curve; larger values are better}
+    \caption{Speedup as core count increases (measured against the single-core run) on the 4890 with a fixed resolution ($400\times400$); the orange line is the theoretical maximum speedup, given linear scaling and 50 execution units (800 cores / 16 cores per unit); the red line is a linear speedup curve; larger values are better}
     \label{fig:strongPlotTwo}
@@
 
+One oft-cited measure of the performance of a computational device is its {\it performance-per-dollar}; this is especially of interest for consumers with budgets or people building large computation clusters.
+In order to measure performance-per-dollar, we have to define {\it performance} in the context of our problem.
+In all of the plots in this section, performance is measured as the number of samples per second.
+Figures \ref{fig:costEffectiveness}, \ref{fig:transistorPlot}, and \ref{fig:wattPlot} all use a 500$\times$500 image, with 2000 samples per pixel, for $500,000,000$ total samples.
+
 \section{Future Work}

### 2010-05-06 paper/paper.tex
Starting on future work section

@@
 
-\usepackage[scaled=.92]{helvet}
-\usepackage{times}
 \usepackage{parskip}
@@
 
-\title{Plotting the Single-Electron Solution to Schr\"{o}dinger's Equation with OpenCL}
+\title{\Large Plotting the Single-Electron Solution to Schr\"{o}dinger's Equation with OpenCL}
 
@@
 
-\subsection{Spherical Harmonics\label{sphericalHarmonics}}
+\subsection{Spherical Harmonics}
 
-Gotta write a blurb here:
+\label{sphericalHarmonics}
+
+$Y$, a solution to Laplace's spherical harmonic function, provides the angular component of $\psi$:
 
@@
 
-\subsection{Laguerre and Legendre\label{laguerreLegendre}}
+\subsection{Laguerre and Legendre}
+
+\label{laguerreLegendre}
 
@@
 
+Each benchmark point is the minimum of five trials.
+
 \subsection{Strong Scaling}
@@
 
+Figure \ref{fig:costEffectiveness} illustrates the performance-per-dollar of our four devices.
+As is quite evident already, the 4890 --- though in the same price range as most of the other devices, and significantly cheaper than the Core 2 Quad --- manages to dominate the chart.
+It's safe to say that if an institution is looking into buying a large number of a particular device for scientific computation, they should seriously consider looking into high-end graphics cards.
+Indeed; each dollar spent on a 4890 goes 18 times farther, performance-wise, than it would if spent on the next-most-cost-effective chip, the Core 2 Duo E7200.
+
+Another important point when considering the cost of various pieces of hardware is how much it will cost to run.
+One way to measure this is to consider the power efficiency of the chip, in terms of performance-per-watt.
+Again using the same measure of performance from our performance-per-dollar comparison, we can see in figure \ref{fig:wattPlot} that the GPU continues to dominate, even though it requires three times more energy to run than its closest competitor, the Core 2 Quad.
+As one might expect, the older --- thus, less power efficient --- Core 2 Duo falls behind the other CPUs on this scale: though it was the performance-per-dollar winner out of the three CPUs, it would cost significantly more than the other two to run for any period of time.
+
 \section{Future Work}
+
+There are quite a few things which have come up while developing this project that would be interesting to implement if we were to continue work on the codebase.
+
+\subsection{Camera Transformation}
+
+\label{cameraTransformation}
+
+One key feature that is missing from the current implementation is the ability to manipulate the "camera" location, changing your viewport on the orbital cloud.
+Implementing this would add a bit of complexity, but would allow for interesting video rendering --- for example, video of the camera spinning around the atom, providing a better idea of the actual shape of the cloud.
+It can be hard --- without prior knowledge --- to comprehend the shape of the cloud from our current visualization.
+
+In addition, if the number of samples was sufficiently reduced, or a fast enough video card was obtained, one could implement {\it live} manipulation of the camera's view, which could assist in inspection of the orbital cloud to an even greater extent.
+
+\subsection{Slicing}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{320-slice.png}
+    \caption{A slice through the 3-2-0 orbital of the hydrogen atom}
+    \label{fig:slice}
+\end{figure}
+
+Section \ref{cameraTransformation} covered the ideal solution to the problem of comprehending the shape of the orbital from our 2D representation --- however, if that solution proved to be too challenging, one could instead generate an animation of slices through the cloud, each slice being similar to figure \ref{fig:slice}
+
+\subsection{Coloring}
+
+\subsection{Generalization of Parameters}
 

### 2010-05-06 paper/paper.tex
More in the future work section

@@
 
-The generating differential equations for these polynomials are as follows:
+The generating equations for these polynomials are as follows:
 
@@
 
-where $\tilde{F}$ is Gauss' hypergeometric function, which we don't care about since we're simply allowing Mathematica to evaluate it and using the comparatively simple symbolic results in our table.
+where $\tilde{F}$ is Gauss' hypergeometric function, which we don't need to bother ourselves with since we're simply allowing Mathematica to evaluate it and using the comparatively simple symbolic results in our table.
 
@@
 
-Each benchmark point is the minimum of five trials.
+Three computers have spent countless hours performing calculations to bring you the following results, as each benchmark point is the minimum of five individual trials.
 
@@
 For this benchmark, we will fix the image at $400\times400$ pixels, requiring $320,000,000$ total samples, and we will vary the number of cores of the 4890 that we use.
+
+One can see in figure \ref{fig:strongPlotTwo} that this implementation is strongly scalable; the speedup is almost perfectly linear as the core count increases.
+It should be noted that since the 4890 only has a maximum of 50 compute units, the linear increase is only maintained until we reach 50 cores; after that, the performance gains plateau as they should.
 
@@
 
-One can see in figure \ref{fig:strongPlotTwo} that this implementation is strongly scalable; the speedup is almost perfectly linear as the core count increases.
-It should be noted that since the 4890 only has a maximum of 50 compute units, the linear increase is only maintained until we reach 50 cores; after that, the performance gains plateau as they should.
-
 \subsection{Large- vs. Small-scale Parallelism}
@@
 
+As one can see in figure \ref{fig:runtimePlot}, the GPU significantly outperforms all of the CPUs.
+This is not surprising, as it can perform more than an order of magnitude more parallel computations, but it does validate the premise of this experiment.
+
 \begin{figure}
@@
 \end{figure}
-
-As one can see in figure \ref{fig:runtimePlot}, the GPU significantly outperforms all of the CPUs.
-This is not surprising, as it can perform more than an order of magnitude more parallel computations, but it does validate the premise of this experiment.
 
@@
 
-There are quite a few things which have come up while developing this project that would be interesting to implement if we were to continue work on the codebase.
+There are quite a few things which have come up while developing this project that would be interesting to implement if we were to continue work on the codebase. We will discuss a few of these below:
 
@@
 
-Section \ref{cameraTransformation} covered the ideal solution to the problem of comprehending the shape of the orbital from our 2D representation --- however, if that solution proved to be too challenging, one could instead generate an animation of slices through the cloud, each slice being similar to figure \ref{fig:slice}
+Section \ref{cameraTransformation} covered the ideal solution to the problem of comprehending the shape of the orbital from our 2D representation --- however, if that solution proved to be too challenging, one could instead generate an animation of slices through the cloud, each slice being similar to figure \ref{fig:slice}.
+This would provide another form of interesting visualization of the resulting data.
 
@@
 
+A third way to make the visualization more appealing (or even more informative) would be to provide some manner of colorization.
+One potential algorithm would be to colorize each sample based on its depth in $z$; the resulting image would then have more depth information, which might increase the ease with which it is interpreted.
+A variety of other coloring functions are possible, but none were pursued during this project, for time's sake.
+
 \subsection{Generalization of Parameters}
+
+Equations \ref{laguerre} and \ref{legendre} provide a generalized way to construct Laguerre and Legendre polynomials, respectively.
+However, our implementation does not make use of these functions, instead using a table of simplified solutions, generated with Mathematica, which is limited in scope.
+
+The length of the table determines the range of $n, l, m$ values which can be passed into our program, and is the only restriction on generality within our implementation.
+If one were to construct these functions on the fly from their generating equations, the algorithm would be completely generalized in terms of $n, l, m$, and would be significantly more useful from an exploratory standpoint.
+
+However, as discussed in section \ref{laguerreLegendre}, this would require (at least) the implementation of a symbolic differential equation solver in OpenCL, which could potentially be a very time-consuming task (it could also increase the number of GPU stream processors per kernel required, significantly decreasing the speedup gained).
+
+Alternatively, one could write a translation program to automatically generate the table from Mathematica, removing the slow, error-prone human translation step, and very quickly expand the size of the table to something more useful.
 

### 2010-05-06 paper/paper.tex
print both computation and copy-to-system-memory times

@@
 
+\subsection{Communication Overhead}
+
+The algorithm we implemented for this project required no communication between  cores whatsoever during the course of the computation.
+The primary overhead involved in the entire process is the single copy of the $n\times n$ output buffer from video memory into system memory, where it is then normalized, compressed, and output (we don't consider these parts of the process when measuring performance or overhead, as they're written in Python, a language not known for its performance characteristics).
+
+The
+
 \subsection{Large- vs. Small-scale Parallelism}

### 2010-05-06 hydrogen.py

@@ -37,13 +37,14 @@ def renderOrbitals((ni, li, mi), imageResolution):
 
     # Evaluate the rest of psi once for each pixel, copy into output buffer
     before = time()
-    main.prg.density(main.queue, [pointCount], ipsi, n, l, m, outputBuffer, res)
+    main.prg.density(main.queue, [pointCount], ipsi, n, l, m, outputBuffer, res).wait()
+    computeDuration = time() - before
     cl.enqueue_read_buffer(main.queue, outputBuffer, output).wait()
-    duration = time() - before
+    copyDuration = time() - before - computeDuration
 
     outputBuffer.release()
 
-    return (output, duration)
+    return (output, (computeDuration, copyDuration))
 
 def exportImage(output):
     """Export a visual representation of the computed density of the
@@ -63,18 +64,20 @@ def exportImage(output):
 def benchmark(skip, params=(3,2,0)):
     """Time computation of the image at various different resolutions"""
     for res in range(100,1010,skip):
-        minDuration = 100000000
+        minComputeDuration = 100000000
+        minCopyDuration = 100000000
 
         for itr in range(5):
-            (output, duration) = renderOrbitals(params, res)
+            (output, (computeDur, copyDur)) = renderOrbitals(params, res)
             if max(output) > 0.0:
-                minDuration = min(minDuration, duration)
+                minComputeDuration = min(minComputeDuration, computeDur)
+                minCopyDuration = min(minCopyDuration, copyDur)
 
         # At some resolutions, on my mobile GPU, I get a totally black image
         # I have no idea why this happens, but this sits here to discard
         # those results, since their timing seems to be somewhat inaccurate.
-        if minDuration:
-            print "{0},{1}".format(res, minDuration)
+        if minComputeDuration < 100000000:
+            print "{0},{1},{2}".format(res, minComputeDuration, minCopyDuration)
 
 def main():
     # Parse commandline arguments
@@ -128,12 +131,12 @@ def main():
     elif options.longBenchmark:
         benchmark(10, params=params)
     elif options.onebench:
-        (output, duration) = renderOrbitals(params, int(options.res))
-        print "{0},{1}".format(options.res, duration)
+        (output, (computeD, copyD)) = renderOrbitals(params, int(options.res))
+        print "{0},{1},{2}".format(options.res, computeD, copyD)
     else:
-        (output, duration) = renderOrbitals(params, int(options.res))
[…]

### 2010-05-06 paper/paper.tex
More plots
Clear up some things

--- 2010-05-06 paper/paper.tex
@@
 
-Once computation has completed, the buffer is copied back from video memory to system memory (or, if the program was running on the CPU, this step is skipped --- this is the primary overhead incurred by the use of GPU computation).
-There, pixel values are scaled linearly so that the brightest pixel is 100\% white.
+Once computation has completed, the buffer is copied back from working memory (either video memory or system memory) to the final image buffer in system memory.
+Once there, pixel values are scaled linearly so that the brightest pixel is 100\% white.
 
@@
 One can see in figure \ref{fig:strongPlotTwo} that this implementation is strongly scalable; the speedup is almost perfectly linear as the core count increases.
-It should be noted that since the 4890 only has a maximum of 50 compute units, the linear increase is only maintained until we reach 50 cores; after that, the performance gains plateau as they should.
+It should be noted that since --- for our kernel --- the 4890 only has 50 compute units, the linear increase is only maintained until we reach 50 cores; after that, the performance gains plateau as they should.
 
@@
 \end{figure}
+
+\subsection{Communication Overhead}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{overheadPlot.pdf}
+    \caption{Overhead from image buffer copy as a percentage of the total runtime, lower numbers are better}
+    \label{fig:overheadPlot}
+\end{figure}
+
+The algorithm we implemented for this project required no communication between  cores whatsoever during the course of the computation.
+The primary overhead involved in the entire process is the single copy of the $n\times n$ output buffer from working memory into system memory, where it is then normalized, compressed, and output (we don't consider these parts of the process when measuring performance or overhead, as they're written in Python, a language not known for its performance characteristics).
+
+The overhead incurred during this copy can be seen in the chart in figure \ref{fig:overheadPlot} to be negligible --- in the range of a ten-thousandth of a percent of the total runtime when working on the GPU, and a one-hundred-thousandth of a percent when working on the CPU, where the data has a shorter distance to travel.
 
@@
     \label{fig:costEffectiveness}
-\end{figure}
-
-\begin{figure}
-    \includegraphics[width=84.5mm]{transistorPlot.pdf}
-    \caption{Samples per second per million transistors with a fixed resolution ($500\times500$); larger values are better}
-    \label{fig:transistorPlot}
 \end{figure}

### 2010-05-07 paper/paper.tex
Update plots, write more words

@@
 
-\section{Introduction}
-
-asdfasdfadsf asdf adsf asdf asdf adfasdf adsf adf asfd
-
-asdfasdfadsf asdf adsf asdf asdf adfasdf adsf adf asfd
-
-asdfasdfadsf asdf adsf asdf asdf adfasdf adsf adf asfd
-
-asdfasdfadsf asdf adsf asdf asdf adfasdf adsf adf asfd
-
 \section{Physics}
@@
 
+This is the fundamental formula involved in our computation --- it is this equation which we evaluate at each sample.
+The result of this equation gives the probability that the single electron of our hydrogen atom is at the particular chosen location in the spherical coordinate system, $\left(\theta, \phi, r\right)$, which is the information that we eventually aim to plot.
+
 One will notice that $\psi_c$ is independent of the coordinates of evaluation, $\left(\theta, \phi, r\right)$, so it --- as an obvious optimization --- can be computed once and cached for the entire image.
 
-Also, $\psi$ depends on two external functions: $Y$, the spherical harmonic equation (equation \ref{yFunc}), and $L$, the Laguerre polynomial (equation \ref{laguerre}).
+Also, $\psi$ depends on two external functions: $Y$, the spherical harmonic equation (equation \ref{yFunc}), and $L$, the Laguerre polynomial (equation \ref{laguerre}), which will be detailed in sections \ref{sphericalHarmonics} and \ref{laguerreLegendre}, respectively.
 
@@
 
-The overhead incurred during this copy can be seen in the chart in figure \ref{fig:overheadPlot} to be negligible --- in the range of a ten-thousandth of a percent of the total runtime when working on the GPU, and a one-hundred-thousandth of a percent when working on the CPU, where the data has a shorter distance to travel.
+The overhead incurred during this copy can be seen in the chart in figure \ref{fig:overheadPlot} to be negligible --- in the range of a tenth of a percent of the total runtime when working on the GPU, and a one-hundred-thousandth of a percent when working on the CPU, where the data has a shorter distance to travel.
 
@@
 In all of the plots in this section, performance is measured as the number of samples per second.
-Figures \ref{fig:costEffectiveness}, \ref{fig:transistorPlot}, and \ref{fig:wattPlot} all use a 500$\times$500 image, with 2000 samples per pixel, for $500,000,000$ total samples.
+Figures \ref{fig:costEffectiveness} and \ref{fig:wattPlot} all use a 500$\times$500 image, with 2000 samples per pixel, for $500,000,000$ total samples.
 
@@
 Section \ref{cameraTransformation} covered the ideal solution to the problem of comprehending the shape of the orbital from our 2D representation --- however, if that solution proved to be too challenging, one could instead generate an animation of slices through the cloud, each slice being similar to figure \ref{fig:slice}.
-This would provide another form of interesting visualization of the resulting data.
+This would provide another form of interesting visualization of the resulting data, rather similar to the data collected by an MRI device.
+
+\subsection{Automatic Scaling}
+
+Our OpenCL kernel currently contains an arbitrary zoom factor which needs to be adjusted based on the orbital configuration --- higher energy electron states lead to larger probability clouds.
+Automatically generating this zoom factor would lead to significantly improved ease of use when changing between different configurations.
+This could most effectively be implemented by finding a formula for a reasonable scale factor given a set of $n, l, m$ values.
 
@@
 
+After consuming many computer-hours computing the 3-2-0 orbital repeatedly, I've come to a few clear conclusions.
+Firstly, scientific computation --- the calculation of atomic orbitals, at the very least --- should clearly be a prime target for
+
 \section{Code}

### 2010-05-07 hydrogen.py
Remove computation of independent part of psi. Speedup near 10%. This was just being used as a scaling term, so it doesn't actually matter for making our plots.

@@ -9,13 +9,6 @@ from time import time
 from PIL import Image
 from optparse import OptionParser
 
-def independentPsi(n, l):
-    """The first term of psi is constant with respect to the entire image, so
-    it can be computed outside of the inner loop."""
-    rootFirst = (2.0 / n) ** 3
-    rootSecond = factorial(n - l - 1.0) / (2.0 * ((n * factorial(n + l)) ** 3))
-    return sqrt(rootFirst * rootSecond)
-
 def renderOrbitals((ni, li, mi), imageResolution):
     """Create various buffers to shuttle data to/from the CPU/GPU, then
     execute the kernel for each pixel in the input image, and copy the
@@ -32,12 +25,9 @@ def renderOrbitals((ni, li, mi), imageResolution):
     outputBuffer = cl.Buffer(main.ctx, mf.READ_WRITE | mf.COPY_HOST_PTR,
                              hostbuf=output)
 
-    # Evaluate first term of psi once for entire image
-    ipsi = numpy.float32(independentPsi(n, l))
-
     # Evaluate the rest of psi once for each pixel, copy into output buffer
     before = time()
-    main.prg.density(main.queue, [pointCount], ipsi, n, l, m, outputBuffer, res).wait()
+    main.prg.density(main.queue, [pointCount], n, l, m, outputBuffer, res).wait()
     computeDuration = time() - before
     cl.enqueue_read_buffer(main.queue, outputBuffer, output).wait()
     copyDuration = time() - before - computeDuration

### 2010-05-07 paper/paper.tex
More workds

@@
 
+\subsection{Multiple Electrons}
+
+Another feature that would be incredibly cool to implement would be simulation of atoms with more than one electron.
+In this case, interference between the electrons changes the pattern to be significantly more complex (and much more interesting, as well).
+It's also a good bit more complex to compute: there is no exact solution of the Schr\"{o}dinger's equation for multiple electrons.
+Instead, an implementation would depend on a numerical differential equation solver, which would significantly affect performance.
+
 \section{Conclusion}
 
-After consuming many computer-hours computing the 3-2-0 orbital repeatedly, I've come to a few clear conclusions.
-Firstly, scientific computation --- the calculation of atomic orbitals, at the very least --- should clearly be a prime target for
+After consuming many computer-hours computing the 3-2-0 orbital repeatedly, we've come to a few clear conclusions.
+Firstly, scientific computation --- the calculation of atomic orbitals, at the very least --- should clearly be a prime target for porting to parallel computation on the GPU, as the field takes off.
+Projects like Folding@Home are already taking advantage of this (besides being massively parallel, as a web-distributed project) by providing ATI Stream SDK and CUDA ports of their client, and the last few years have seen an explosion in other such projects.
+
+We've also found that --- for problems which are embarrassingly easy to parallelize, like this one --- it's very easy to implement your problem using OpenCL.
+We've used MPI and pthreads in the past, and it seems that OpenCL (perhaps because it's a more modern API) is a bit easier to use in terms of implementation.
+The added benefit of kernels running on a wider array of hardware (not just the CPU, like MPI, but on the GPU as well) means that OpenCL is a no-brainer --- at least for this sort of simple project.
+It's quite likely that a problem requiring a large amount of inter-kernel communication would be better suited by something with more communications primitives, like MPI.
 

### 2010-05-08 paper/paper.tex
Paper

@@
 
+Quantum mechanics --- since Louis de Broglie's paper {\it Research on Quantum Theory} --- tells us that matter behaves sometimes as a wave, and other times as a particle.
+The electron in a hydrogen atom is no different: when it isn't being observed, it doesn't exist at a particular point in space, but instead expands as a wave.
+The atomic orbital cloud which we're plotting here is actually simply a plot of the density of that wave at all points surrounding the atom's nucleus.
+
 All of the equations below are borrowed from \cite{quantumBook}, but have been slightly modified to fit our purposes, as discussed in section \ref{simpSection}.
@@
 One will notice that $\psi_c$ is independent of the coordinates of evaluation, $\left(\theta, \phi, r\right)$, so it --- as an obvious optimization --- can be computed once and cached for the entire image.
+
+Indeed, while all of the benchmarks in this paper were done with $\psi_c$ included, one might notice that --- since it's a constant term across the entire image, and we're scaling the image by {\it another} constant scale factor --- that it doesn't actually affect the visualization, and can be entirely discarded.
+It appears that discarding the overhead required to pass this value in as an argument to each of the kernel instances and to multiply it into the value of $\psi$ gives a performance increase of between 6 and 10 percent: quite a significant improvement for no change in output!
 
@@
 Once computation has completed, the buffer is copied back from working memory (either video memory or system memory) to the final image buffer in system memory.
-Once there, pixel values are scaled linearly so that the brightest pixel is 100\% white.
+Once there, pixel values are scaled linearly so that the brightest pixel is 100\% white, and the image is compressed to PNG and saved to the filesystem.
 
@@
 
+\subsection{Communication Overhead}
+
+\begin{figure}
+    \includegraphics[width=84.5mm]{overheadPlot.pdf}
+    \caption{Overhead from image buffer copy as a percentage of the total runtime; smaller values are better}
+    \label{fig:overheadPlot}
+\end{figure}
+
+The algorithm we implemented for this project required no communication between  cores whatsoever during the course of the computation.
+The primary overhead involved in the entire process is the single copy of the $n\times n$ output buffer from working memory into system memory, where it is then normalized, compressed, and output (we don't consider these parts of the process when measuring performance or overhead, as they're written in Python, a language not known for its performance characteristics).
+
+The overhead incurred during this copy can be seen in the chart in figure \ref{fig:overheadPlot} to be negligible --- in the range of a tenth of a percent of the total runtime when working on the GPU, and a one-hundred-thousandth of a percent when working on the CPU, where the data has a shorter distance to travel.
+
 \subsection{Strong Scaling}
@@
 \label{strongScaling}
-
-In order to measure the strong scalability of our implementation, we fix the problem size and vary the number of processing units used to compute it.
-For this benchmark, we will fix the image at $400\times400$ pixels, requiring $320,000,000$ total samples, and we will vary the number of cores of the 4890 that we use.
-
-One can see in figure \ref{fig:strongPlotTwo} that this implementation is strongly scalable; the speedup is almost perfectly linear as the core count increases.
-It should be noted that since --- for our kernel --- the 4890 only has 50 compute units, the linear increase is only maintained until we reach 50 cores; after that, the performance gains plateau as they should.
 
@@
 
-\subsection{Communication Overhead}
+In order to measure the strong scalability of our implementation, we fix the problem size and vary the number of processing units used to compute it.
+For this benchmark, we will fix the image at $400\times400$ pixels, requiring $320,000,000$ total samples, and we will vary the number of cores of the 4890 that we use.
 
-\begin{figure}
-    \includegraphics[width=84.5mm]{overheadPlot.pdf}
-    \caption{Overhead from image buffer copy as a percentage of the total runtime, lower numbers are better}
-    \label{fig:overheadPlot}
-\end{figure}
-
-The algorithm we implemented for this project required no communication between  cores whatsoever during the course of the computation.
-The primary overhead involved in the entire process is the single copy of the $n\times n$ output buffer from working memory into system memory, where it is then normalized, compressed, and output (we don't consider these parts of the process when measuring performance or overhead, as they're written in Python, a language not known for its performance characteristics).
-
-The overhead incurred during this copy can be seen in the chart in figure \ref{fig:overheadPlot} to be negligible --- in the range of a tenth of a percent of the total runtime when working on the GPU, and a one-hundred-thousandth of a percent when working on the CPU, where the data has a shorter distance to travel.
+One can see in figure \ref{fig:strongPlotTwo} that this implementation is strongly scalable; the speedup is almost perfectly linear as the core count increases.
+It should be noted that since --- for our kernel --- the 4890 only has 50 compute units, the linear increase is only maintained until we reach 50 cores; after that, the performance gains plateau as they should.
 
@@
 It's quite likely that a problem requiring a large amount of inter-kernel communication would be better suited by something with more communications primitives, like MPI.
+Indeed, besides {\it barrier}, OpenCL doesn't seem to have any manner of inter-kernel communications functions.
 
